Architectures and Open pNets

### 2018-10-11 2018-Arch-pNets/arch-pNets.tex
Arch-pNets paper: updated the architecture figure; minor edits

\newcommand{\ordbool}{\ensuremath{\sB^{<}}}
\newcommand{\data}{\ensuremath{\sD}}
\newcommand{\signature}{\ensuremath{\Sigma}}
\newcommand{\variables}{\ensuremath{\cV}}
\newcommand{\Talg}{\ensuremath{\cT_{\signature,\variables}}}
\newcommand{\actions}[1]{\ensuremath{\cA_{#1}}}
\newcommand{\guards}[1]{\ensuremath{\ordbool_{#1}}}
\newcommand{\exprs}[1]{\ensuremath{\sE_{#1}}}
\newcommand{\boolexprs}[1]{\ensuremath{\sB_{#1}}}
\newcommand{\assigns}[1]{\ensuremath{\sA_{#1}}}
\newcommand{\valuations}[1]{\ensuremath{\data^{#1}}}
\newcommand{\val}[3][]{\ensuremath{#1{\sigma}^{#2}_{#3}}}

@@

\newcommand{\primeit}[1]{#1'}
\newcommand{\doubleprimeit}[1]{#1''}

\newcommand{\export}[1][]{\ensuremath{\varepsilon_{#1}}}
\newcommand{\valdiff}[2]{\ensuremath{#1 \triangle #2}}
\newcommand{\supp}[1]{\ensuremath{\mathrm{supp}(#1)}}
\newcommand{\semopen}[1]{\ensuremath{[{#1}]}}
\newcommand{\semclosed}[1]{\ensuremath{\llbracket{#1}\rrbracket}}
\newcommand{\reachable}[1]{\ensuremath{\mathit{reachable}({#1})}}
\newcommand{\IMextend}[2]{\ensuremath{#1 \ltimes #2}}
\newcommand{\arcomp}{\oplus}
\newcommand{\arequiv}{\equiv}
\newcommand{\Arcomp}{\Bigoplus}
\newcommand{\expmix}{\wedge}
\newcommand{\order}{\leqslant}

@@

  \begin{tabular}{
      @{}p{0.15\columnwidth}@{}
      p{0.35\columnwidth}@{}
      p{0.5\columnwidth}@{}
    }
    {\bf Output} & {\bf Command} & {\bf Definition}
    \\
    $\data$ &
    {\ttfamily\textbackslash data} &
    {\ttfamily\textbackslash sD}
    \\
    \signature &
    {\ttfamily\textbackslash signature} &
    {\ttfamily\textbackslash Sigma}
    \\
    \variables &
    {\ttfamily\textbackslash variables} &
    {\ttfamily\textbackslash cV}
    \\
    \Talg &
    {\ttfamily\textbackslash Talg} &
    {\ttfamily\textbackslash cT\_\{\textbackslash signature,\textbackslash variables\}}
    \\
    \actions{\#1} &
    {\ttfamily\textbackslash actions\{\#1\}} &
    {\ttfamily\textbackslash cA\_\{\#1\}}
    \\
    \exprs{\#1} &       
    {\ttfamily\textbackslash exprs\{\#1\}} &
    {\ttfamily\textbackslash sE\_\{\#1\}}
    \\
    \boolexprs{\#1} &       
    {\ttfamily\textbackslash boolexprs\{\#1\}} &
    {\ttfamily\textbackslash sB\_\{\#1\}}
    \\
    \guards{\#1} &
    {\ttfamily\textbackslash guards\{\#1\}} &
    {\ttfamily\textbackslash sB\^{}\{<\}\_\{\#1\}}
    \\
    \assigns{\#1} &
    {\ttfamily\textbackslash assigns\{\#1\}} &
    {\ttfamily\textbackslash sA\_\{\#1\}}
    \\
    \valuations{\#1} &
    {\ttfamily\textbackslash valuations\{\#1\}} &
    {\ttfamily\textbackslash data\^{}\{\#1\}}
    \\[6pt]
    \val[]{\#2}{\#3} &
    {\ttfamily\textbackslash val[]\{\#2\}\{\#3\}} &
    {\ttfamily\#1\{\textbackslash sigma\}\^{}\{\#2\}\_\{\#3\}}
    \\[6pt]
    \val[\tilde]{\#2}{\#3} &
    {\ttfamily\textbackslash val[\textbackslash tilde]\{\#2\}\{\#3\}} &
    {\ttfamily\#1\{\textbackslash sigma\}\^{}\{\#2\}\_\{\#3\}}
  \end{tabular}
\noteSBin{commands above (tabular generates an error inside a
  note)}

Formally, we assume the existence of a term algebra $\Talg$,
where $\signature$ is the signature of the data and action constructors,
and $\variables$ a set of \emph{variables}. Within $\Talg$, we distinguish a set of
\emph{data expressions} $\exprs{\variables}$ \replaceSB{, including}{and} a set of \emph{Boolean
expressions} \todoLH{$\mathcal{B} \to g$}\todoSB{change to $\boolexprs{\variables}$?}
$\guards{\variables}$ ($\guards{\variables}\subseteq\exprs{\variables}$).
On top of $\exprs{\variables}$ we build the \emph{action algebra}
$\actions{\variables}$, with $\actions{\variables}\subseteq\Talg$ \replaceSB{,}{and}
$\exprs{\variables}\cap\actions{\variables} = \emptyset$;
naturally action terms will use data expressions as subterms.
\addSB{In addition to the above notations, we will use
\[
\assigns{\variables} \bydef{=}
\Setdef{(x_i := e_i)^{i \in I}}{
  x_i^{i \in I}\subseteq \variables,
  e_i^{i \in I}\subseteq \exprs{\variables},
  I \text{ is a finite index set}
}
\]
to denote the set of \emph{variable assignments}.}

@@

\addSB{%
%
  Finally, we assume the existence of a universal data domain
  given as a partially-ordered set $(\data, \order)$, potentially
  encompassing several copies of any given data type with
  different orders.  For example, we assume that $(\data,
  \order)$ comprises both the unordered set of Booleans $\sB =
  (\{\true, \false\}, \emptyset)$ and the naturally ordered one
  $\sB^< = (\{\true, \false\}, \{\false < \true\})$, and
  similarly for integer and real numbers; as well as the set of
  intervals ordered by inclusion.

  For a set of variables $V \subseteq \variables$, we denote
  $\valuations{V} \bydef{=} \{\val{}{}: V \rightarrow \data\}$
  the set of \emph{valuations} of the variables in $V$.
  Valuations extend canonically to expressions.  Thus
  $\val{}{}(x)$ and $\val{}{}(e)$ denote, respectively, the
  values of the variable $x$ and the expression $e$ under the
  valuation $\val{}{}$.  For a valuation $\val{}{} \in
  \valuations{V}$ and an assignment $(x_i := e_i)^{i \in I} \in
  \assigns{V}$, we denote $\val{}{}[(x_i := e_i)^{i \in I}]$ the
  valuation defined by
  \[
  \val{}{}\bigl[(x_i := e_i)^{i \in I}\bigr](x) \bydef{=}
  \begin{cases}
    \val{}{}(x), & \text{if } x \not\in x_i^{i \in I}\,,\\
    \val{}{}(e_i), & \text{if } x = x_i, \text{for some } i \in I\,.
  \end{cases}
  \]
  For two valuations $\val{1}{}, \val{2}{} : V \rightarrow
  \data$, we denote \[\valdiff{\val{1}{}}{\val{2}{}} \bydef{=}
  \setdefb{v \in V}{\val{1}{}(v) \neq \val{2}{}(v)}\] the set of
  variables that are assigned different values by the two
  valuations.  As usual, we write $\val{1}{} \order \val{2}{}$
  iff $\val{1}{}(x) \order \val{2}{}(x)$, for all $x \in V$.  An
  expression $e \in \exprs{V}$ is called \emph{monotonic} if, for
  any two valuations $\val{1}{}, \val{2}{} \in \valuations{V}$,
  $\val{1}{} \order \val{2}{}$ implies $\val{1}{}(e) \order
  \val{2}{}(e)$.
%
}

@@

\begin{definition}[Components]
  \label{defn:component}
  A \emph{component} is a tuple $(Q, q^0, V, \val{0}{}, P,
  \export, \goesto{})$, where
  \begin{itemize}
  \item $Q$ is a set of \emph{states}, with $q^0 \in Q$ the
    \emph{initial state}, 
  \item $V$ is a set of \emph{component variables},
  \item $\val{0}{} : V \rightarrow \data$ is an \emph{initial
    valuation} of the component variables, 
  \item $P$ is a set of \emph{ports}, with
    \noteSB{Added this}\hl{$\export : P \rightarrow 2^V$ 
    the \emph{set of variables exported by each port},}
  \item $\goesto{}\, \subseteq
    Q \times 2^P \times \guards{V} \times \exprs{V} \times Q$
%
    is a \emph{transition relation}, with transitions
    labelled by \emph{interactions}, \ie triples consisting of a non-empty set of ports
    $\emptyset \neq a \subseteq P$, a monotonic Boolean \emph{guard} $g \in
    \guards{V}$ and a monotonic \emph{update expression} \todoSB{$e \rightarrow u$}\hl{$e \in
    \exprs{V}$}.
  \end{itemize}
%
  We call the triple $(V,P, \export)$ the \emph{interface} of the
  component.\,\footnote{%
%
  In practice, component variables are split in two sub-sets:
  local and exported variables.  Only exported variables belong
  to the component interface and can be used for the interactions
  with other components (see \defn{im}).  However, in the context
  of this paper we can simplify by omitting this separation.
%
  }
%
  Notations $q \goesto{a, g, e} q'$ and $q \goesto{a, g, e}$ are as usual;
  for a component $B$, we denote $Q_B$, $q^0_B$, $V_B$, $\val{0}{B}$, $P_B$,
  and $\export[B]$ the corresponding constituents of $B$.  We will
  skip the index on the transition relations $\goesto{}$, since it
  is always clear from the context.
\end{definition}

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{BIPspec-ArchFailureTimerMax-v3}
  \caption{The BIP specification of the Failure Monitor architecture}
  \label{fig:schema:ArchFailure:BIP}
\end{figure}

@@

The model in \fig{schema:ArchFailure:BIP} has three components.  The
component $\NameOprnd$ is an \emph{operand} of the architecture, \ie
it only has its interface defined: $(\emptyset, \{\PortFinish,
\PortResume, \PortFail\}, \export[\emptyset])$, with
$\export[\emptyset](p) = \emptyset$, for all $p \in \{\PortFinish,
\PortResume, \PortFail\}$.  The other two components, $\NameCtrl$ and
$\NameTimer$, are \emph{coordinators}\mdash they are specified
completely, including the states and transitions defining their
\emph{behaviour}.  Transitions are labelled with the ports of the
corresponding components, Boolean guards and update functions on local
variables.  For instance the loop transition $\tstate{1}
\goesto{\PortTick, \GuardTick[], \TimerUpd[]} \tstate{1}$ in the
$\NameTimer$ component is labeled by the port $\PortTick$, it can be
fired only when the current value of the local variable $t$ is less
than $z.u$, where $z = [l, u]$ is a local variable representing an
interval.  Upon firing, this transition decrements the value of $t$ by
$1$.  The guard and update function of the transition $\tstate{1}
\goesto{\PortResume} \tstate{0}$ are omitted.  The default values are
the constant predicate $\true$ and the $\mathit{skip}$ operator
respectively.  The initial valuations $\val{0}{\NameTimer} = \{z
\mapsto [0,0], t \mapsto 0\}$ and $\val{0}{\NameCtrl} =
\{\mathit{zone} \mapsto [\mathrm{Min}, \mathrm{Max}]\}$ are shown by
initialising transitions $\goesto{\TimerInitZone[]} \tstate{0}$ and
$\goesto{\IntervalInit[]} \cstate{0}$ respectively.  The constants
$\mathrm{Min}$ and $\mathrm{Max}$ are the parameters of the
architecture.
%\noteLH{It looks like Max is in zone, I do not understand}

@@

\begin{definition}[Component semantics]
  \label{defn:comp:semantics}
  The \emph{open semantics} of a component $B = (Q, q^0, V,
  \val{0}{}, P, \export, \goesto{})$ is given by the LTS denoted $\semopen{B} = (S,
  s^0, \goesto{})$, where $S = Q \times \valuations{V}$, $s^0 =
  (q^0, \val{0}{})$ and $\goesto{}$ is the minimal transition
  relation satisfying the rule
  %
  \begin{equation}
    \label{eq:comp:semantics}
    \infer{
      q \goesto{a, g, e} q'
      \and
      \val{}{} \models g
      \and
      \val[\primeit]{}{} = \tilde{\val{}{}}[e]
      \and
      \valdiff{\val{}{}}{\tilde{\val{}{}}} \subseteq \export(a)
    }{
      (q, \val{}{}) \goesto{a,\tilde{\val{}{}}} (q', \val[\primeit]{}{})
    }
    \,.
  \end{equation}
  %
  \todoSB{Do we actually need this?}\todoLH{I agree: I do not think we need this}
  \hl{The \emph{closed semantics} of $B$ is given by the LTS denoted
  $\semclosed{B} \bydef{=}\reachable{\semopen{B}}$, comprising
  only the reachable states of $\semopen{B}$.}
\end{definition}

@@

  An \emph{interaction model over \hl{$(V, P, \export)$}} is a set $\Gamma
  \subseteq 2^P \times \guards{V} \times \exprs{V}$, such that,
  for any interaction $(a, g, e) \in \Gamma$, we have
  $g \in \guards{\export(a)}$ and $e \in \exprs{\export(a)}$.\footnote{%
%
    Notice that this definition allows $(\emptyset, \true,
    \noop)$ and $(\emptyset, \false, \noop)$ to be included in
    $\Gamma$.
%
  }
  %% We call the set of ports $P$ the \emph{domain} of the
  %% interaction model.
\end{definition}

%% \noteLH{supp should go outside the definition}
When speaking of a set of components or interfaces , we will always
assume that it satisfies the disjointness assumption of \defn{im}.
%
We call the \emph{support} of a set of ports $a \subseteq P$, denoted
$\supp{a}$, either the set $\setdef{i \in I}{a \cap P_i \neq
  \emptyset}$ (for $P = \bigcup_{i=1}^n P_i$) or the set $\setdef{B
  \in \cB}{a \cap P_B \neq \emptyset}$ (for $P = \bigcup_{B \in \cB}
P_B$).  The precise meaning of this notation will always be clear from
the context.  Intuitively, for an interaction $(a, g, e)$, $\supp{a}$
denotes the set of the participating components.

\begin{definition}[Composition]
  \label{defn:im:sem}
  The \emph{composition} of a finite set of components $\cB = (Q_i,
  q^0_i, V_i, \val{0}{i}, P_i, \export[i], \goesto{})^{i \in I}$ with
  the interaction model $\Gamma$ over $(V, P, \export)$ (as in \defn{im})
  is the component $\Gamma(\cB) = (Q, q^0, V, \val{0}{}, P,
    \export, \goesto{})$, where
%
  $Q = \prod_{i \in I} Q_i$;
%
  $q^0 = (q_i^0)^{i \in I}$;
%
  $\val{0}{}: V \rightarrow \data$ is such that, for any $v \in
  V_i$, $\val{0}{}(v) = \val{0}{i}(v)$;
%
  and $\goesto{}$ is the minimal transition relation satisfying
  the rule
%
  \begin{gather}
%%     \label{eq:im:empty}
%%     \infer{
%%       q_j \goesto{\emptyset, g, e} q_j'
%%       \and
%%       \forall i \neq j, q_i = q_i'
%%     }{
%%       (q_i)^{i \in I} \goesto{\emptyset, g, e} (q_i')^{i \in I}
%%     }\,,
%% %
%%     \\[0.5\baselineskip]
    \label{eq:im:sem}
%
    \infer{    
      (a, g, e) \in \Gamma
      \and
      a \neq \emptyset
      \and
      \forall i \in \supp{a}, q_i \goesto{a \cap P_i, g_i, e_i} q_i'
      \and
      \forall i \not\in \supp{a}, q_i = q_i'
%
      \\\\
      \textstyle
%
      g' = g \land \bigwedge_{i \in \supp{a}} g_i
      \and
      e' = e; e_i^{i \in \supp{a}}
    }{
      (q_i)^{i \in I} \goesto{a, g', e'} (q_i')^{i \in I}
    }\,.
  \end{gather}
  \todoSBin{Add the intuitive explanation}
\end{definition}

@@

For instance, in \fig{schema:ArchFailure:BIP}, the two ports
$\NameTimer.\PortStart$ and $\NameCtrl.\PortFail$ are always
synchronised, since they belong to the same binary sub-connector,
where they are both synchrons.  In particular, this means that
whenever the transition $\cstate{0} \goesto{\PortFail} \cstate{1}$ is
fired, so is the transition $\tstate{0} \goesto{\PortStart,
  \TimerInit[]} \tstate{1}$, initialising the timer.  The binary
connector $\NameTimer.\PortStart \bullet\!\!\!-\!\!\!-\!\!\!\bullet
\NameCtrl.\PortFail$ is a sub-connector of a hierarchical connector,
where the port $\NameOprnd.\PortFail$ is a trigger.  Thus, the above
interaction can only happen together with $\NameOprnd.\PortFail$,
forming a ternary interaction.  On the contrary, being a trigger, the
port $\NameOprnd.\PortFail$ can fire alone, forming a singleton
interaction.

>>>>

%% \newcommand{\order}{\leqslant}
\newcommand{\order}{<}
\newcommand{\ordbool}{\ensuremath{\sB^{\order}}}
\newcommand{\data}{\ensuremath{\sD}}
\newcommand{\signature}{\ensuremath{\Sigma}}
\newcommand{\variables}{\ensuremath{\cV}}
\newcommand{\Talg}{\ensuremath{\cT_{\signature,\variables}}}
\newcommand{\actions}[1]{\ensuremath{\cA_{#1}}}
\newcommand{\exprs}[1]{\ensuremath{\sE_{#1}}}
\newcommand{\monexprs}[1]{\ensuremath{\sE^{\order}_{#1}}}
\newcommand{\boolexprs}[1]{\ensuremath{\sB_{#1}}}
\newcommand{\guards}[1]{\ensuremath{\ordbool_{#1}}}
\newcommand{\assigns}[1]{\ensuremath{\sA_{#1}}}
\newcommand{\updates}[1]{\ensuremath{\sA^{\order}_{#1}}}
\newcommand{\valuations}[1]{\ensuremath{\data^{#1}}}
\newcommand{\val}[3][]{\ensuremath{#1{\sigma}^{#2}_{#3}}}
\newcommand{\emptyassign}{\noop}

@@

\newcommand{\primeit}[1]{#1'}
\newcommand{\secondit}[1]{#1''}

\newcommand{\export}[1][]{\ensuremath{\varepsilon_{#1}}}
\newcommand{\valdiff}[2]{\ensuremath{#1 \triangle #2}}
\newcommand{\supp}[1]{\ensuremath{\mathrm{supp}(#1)}}
\newcommand{\semopen}[1]{\ensuremath{[{#1}]}}
\newcommand{\semclosed}[1]{\ensuremath{\llbracket{#1}\rrbracket}}
\newcommand{\reachable}[1]{\ensuremath{\mathit{reachable}({#1})}}
\newcommand{\IMextend}[2]{\ensuremath{#1 \ltimes #2}}
\newcommand{\arcomp}{\oplus}
\newcommand{\arequiv}{\equiv}
\newcommand{\Arcomp}{\Bigoplus}
\newcommand{\expmix}{\wedge}

@@

  \begin{tabular}{
      @{}p{0.15\columnwidth}@{}
      p{0.35\columnwidth}@{}
      p{0.5\columnwidth}@{}
    }
    {\bf Output} & {\bf Command} & {\bf Definition}
    \\
    $\data$ &
    {\ttfamily\textbackslash data} &
    {\ttfamily\textbackslash sD}
    \\
    \signature &
    {\ttfamily\textbackslash signature} &
    {\ttfamily\textbackslash Sigma}
    \\
    \variables &
    {\ttfamily\textbackslash variables} &
    {\ttfamily\textbackslash cV}
    \\
    \Talg &
    {\ttfamily\textbackslash Talg} &
    {\ttfamily\textbackslash cT\_\{\textbackslash signature,\textbackslash variables\}}
    \\
    \actions{\#1} &
    {\ttfamily\textbackslash actions\{\#1\}} &
    {\ttfamily\textbackslash cA\_\{\#1\}}
    \\
    \exprs{\#1} &       
    {\ttfamily\textbackslash exprs\{\#1\}} &
    {\ttfamily\textbackslash sE\_\{\#1\}}
    \\
    \monexprs{\#1} &       
    {\ttfamily\textbackslash monexprs\{\#1\}} &
    {\ttfamily\textbackslash sE\^{}\{<\}\_\{\#1\}}
    \\
    \boolexprs{\#1} &       
    {\ttfamily\textbackslash boolexprs\{\#1\}} &
    {\ttfamily\textbackslash sB\_\{\#1\}}
    \\
    \guards{\#1} &
    {\ttfamily\textbackslash guards\{\#1\}} &
    {\ttfamily\textbackslash sB\^{}\{<\}\_\{\#1\}}
    \\
    \assigns{\#1} &
    {\ttfamily\textbackslash assigns\{\#1\}} &
    {\ttfamily\textbackslash sA\_\{\#1\}}
    \\
    \updates{\#1} &
    {\ttfamily\textbackslash updates\{\#1\}} &
    {\ttfamily\textbackslash sA\^{}\{<\}\_\{\#1\}}
    \\
    \valuations{\#1} &
    {\ttfamily\textbackslash valuations\{\#1\}} &
    {\ttfamily\textbackslash data\^{}\{\#1\}}
    \\[6pt]
    \val[]{\#2}{\#3} &
    {\ttfamily\textbackslash val[]\{\#2\}\{\#3\}} &
    {\ttfamily\#1\{\textbackslash sigma\}\^{}\{\#2\}\_\{\#3\}}
    \\[6pt]
    \val[\tilde]{\#2}{\#3} &
    {\ttfamily\textbackslash val[\textbackslash tilde]\{\#2\}\{\#3\}} &
    (optional argument usage examples)
    \\[6pt]
    \val[\primeit]{\#2}{\#3} &
    {\ttfamily\textbackslash val[\textbackslash primeit]\{\#2\}\{\#3\}} &
    \\[6pt]
    \val[\secondit]{\#2}{\#3} &
    {\ttfamily\textbackslash val[\textbackslash secondit]\{\#2\}\{\#3\}} &
  \end{tabular}
\noteSBin{commands above (tabular generates an error inside a
  note)}

Formally, we assume the existence of a term algebra $\Talg$,
where $\signature$ is the signature of the data and action constructors,
and $\variables$ a set of \emph{variables}. Within $\Talg$, we distinguish a set of
\emph{data expressions} $\exprs{\variables}$ \replaceSB{, including}{and} a set of \emph{Boolean
expressions} \todoLH{$\mathcal{B} \to g$}
$\boolexprs{\variables}\subseteq\exprs{\variables}$.
On top of $\exprs{\variables}$ we build the \emph{action algebra}
$\actions{\variables}$, with $\actions{\variables}\subseteq\Talg$ \replaceSB{,}{and}
$\exprs{\variables}\cap\actions{\variables} = \emptyset$;
naturally action terms will use data expressions as subterms.
\addSB{In addition to the above notations, we will use
\[
\assigns{\variables} \bydef{=}
\Setdef{(x_i := e_i)^{i \in I}}{
  x_i^{i \in I}\subseteq \variables,
  e_i^{i \in I}\subseteq \exprs{\variables},
  I \text{ is a finite index set}
}
\]
to denote the set of \emph{variable assignments}.}  We use the symbol
$\emptyassign$ to denote the empty assignment, \ie an assignment with
$I = \emptyset$.

@@

Finally, we assume the existence of a universal data domain
given as a partially-ordered set $(\data, \order)$, potentially
encompassing several copies of any given data type with
different orders.  For example, we assume that $(\data,
\order)$ comprises both the unordered set of Booleans $\sB =
(\{\true, \false\}, \emptyset)$ and the naturally ordered one
$\sB^< = (\{\true, \false\}, \{\false < \true\})$, and
similarly for integer and real numbers; as well as the set of
intervals ordered by inclusion.

For a set of variables $V \subseteq \variables$, we denote
$\valuations{V} \bydef{=} \{\val{}{}: V \rightarrow \data\}$
the set of \emph{valuations} of the variables in $V$.
Valuations extend canonically to expressions.  Thus
$\val{}{}(x)$ and $\val{}{}(e)$ denote, respectively, the
values of the variable $x$ and the expression $e$ under the
valuation $\val{}{}$.
%
\addSB{
%
It is important to observe that in practice, and in the examples of
this paper, all the variables are sorted, meaning that their values
can only range over given subsets of $\data$, such as $\sB$ or
$\sB^<$.  Hence, applied to expressions, valuations should be
interpreted as partial functions $\exprs{V} \rightharpoonup
\data$\footnote{
%

%
}
%
}

For a valuation $\val{}{} \in
\valuations{V}$ and an assignment $(x_i := e_i)^{i \in I} \in
\assigns{V}$, we denote $\val{}{}[(x_i := e_i)^{i \in I}]$ the
valuation defined by
\[
\val{}{}\bigl[(x_i := e_i)^{i \in I}\bigr](x) \bydef{=}
\begin{cases}
  \val{}{}(x), & \text{if } x \not\in x_i^{i \in I}\,,\\
  \val{}{}(e_i), & \text{if } x = x_i, \text{for some } i \in I\,.
\end{cases}
\]
For two valuations $\val{1}{}, \val{2}{} : V \rightarrow
\data$, we denote \[\valdiff{\val{1}{}}{\val{2}{}} \bydef{=}
\setdefb{v \in V}{\val{1}{}(v) \neq \val{2}{}(v)}\] the set of
variables that are assigned different values by the two
valuations.  As usual, we write $\val{1}{} \order \val{2}{}$
iff $\val{1}{}(x) \order \val{2}{}(x)$, for all $x \in V$.  An
expression $e \in \exprs{V}$ is called \emph{monotonic} if, for
any two valuations $\val{1}{}, \val{2}{} \in \valuations{V}$,
$\val{1}{} \order \val{2}{}$ implies $\val{1}{}(e) \order
\val{2}{}(e)$.
%
Similarly, an assignment $(x_i := e_i)^{i \in I} \in \assigns{V}$ is
called monotonic if all expressions $e_i^{i \in I}$ are monotonic.
We will denote $\guards{V} \subset \boolexprs{V}$, $\monexprs{V}
\subset \exprs{V}$ and $\updates{V} \subset \assigns{V}$ the sets of
monotonic Boolean and generic expressions and assignments,
respectively.
%

@@

\begin{definition}[Components]
  \label{defn:component}
  A \emph{component} is a tuple $(Q, q^0, V, \val{0}{}, P,
  \export, \goesto{})$, where
  \begin{itemize}
  \item $Q$ is a set of \emph{states}, with $q^0 \in Q$ the
    \emph{initial state}, 
  \item $V$ is a set of \emph{component variables},
  \item $\val{0}{} : V \rightarrow \data$ is an \emph{initial
    valuation} of the component variables, 
  \item $P$ is a set of \emph{ports}, with
    $\export : P \rightarrow 2^V$ 
    the \emph{set of variables exported by each port},
  \item $\goesto{}\, \subseteq
    Q \times 2^P \times \guards{V} \times \updates{V} \times Q$
%
    is a \emph{transition relation}, with transitions
    labelled by \emph{interactions}, \ie triples consisting of a non-empty set of ports
    $\emptyset \neq a \subseteq P$, a monotonic Boolean \emph{guard} $g \in
    \guards{V}$ and a monotonic \emph{update assignment} \hl{$u \in
    \updates{V}$}.
  \end{itemize}
%
  We call the triple $(V,P, \export)$ the \emph{interface} of the
  component.\,\footnote{%
%
  In practice, only exported variables, \ie those in the set
  $\bigcup_{p \in P}\export(p)$, belong to the component interface and
  can be used for the interactions with other components (see
  \defn{im}).  However, in the context of this paper we can simplify
  by omitting this separation.
%
  }
%
  Notations $q \goesto{a, g, u} q'$ and $q \goesto{a, g, u}$ are as usual;
  for a component $B$, we denote $Q_B$, $q^0_B$, $V_B$, $\val{0}{B}$, $P_B$,
  and $\export[B]$ the corresponding constituents of $B$.  We will
  skip the index on the transition relations $\goesto{}$, since it
  is always clear from the context.
\end{definition}

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{BIPspec-ArchFailureTimerMax-v4}
  \caption{The BIP specification of the Failure Monitor architecture}
  \label{fig:schema:ArchFailure:BIP}
\end{figure}

@@

The model in \fig{schema:ArchFailure:BIP} has three components.  The
component $\NameOprnd$ is an \emph{operand} of the architecture, \ie
it only has its interface defined: $(\emptyset, \{\PortFinish,
\PortResume, \PortFail\}, \export[\emptyset])$, with
$\export[\emptyset](p) = \emptyset$, for all $p \in \{\PortFinish,
\PortResume, \PortFail\}$.  The other two components, $\NameCtrl$ and
$\NameTimer$, are \emph{coordinators}\mdash they are specified
completely, including the states and transitions defining their
\emph{behaviour}.  Transitions are labelled with the ports of the
corresponding components, Boolean guards and update assignments on local
variables.  For instance, the loop transition $\tstate{1}
\goesto{\PortTick, \GuardTick[], \TimerUpd[]} \tstate{1}$ in the
$\NameTimer$ component is labeled by the port $\PortTick$, it can be
fired only when the current value of the local variable $t$ is less
than $z.u$, where $z = [l, u]$ is a local variable representing an
interval.  Upon firing, this transition increments the value of $t$ by
$1$.  The guard and update assignment of the transition $\tstate{1}
\goesto{\PortResume} \tstate{0}$ are omitted.  The default values are
the constant predicate $\true$ and the empty assignment {\emptyassign}
%$\mathit{skip}$ operator
respectively.  The initial states \tstate{0} and \cstate{0}, and valuations $\val{0}{\NameTimer} = \{z
\mapsto \topinterval, t \mapsto 0\}$ and $\val{0}{\NameCtrl} =
\{\mathit{zone} \mapsto [\mathrm{Min}, \mathrm{Max}]\}$ are shown by
incoming arrows $\goesto{\TimerInit[]} \tstate{0}$ and
$\goesto{\IntervalInit[]} \cstate{0}$, respectively.  The constants
$\mathrm{Min}$ and $\mathrm{Max}$ are the parameters of the
architecture.
%\noteLH{It looks like Max is in zone, I do not understand}

@@

\begin{definition}[Component semantics]
  \label{defn:comp:semantics}
  The \emph{open semantics} of a component $B = (Q, q^0, V,
  \val{0}{}, P, \export, \goesto{})$ is given by the LTS denoted $\semopen{B} = (S,
  s^0, \goesto{})$, where $S = Q \times \valuations{V}$, $s^0 =
  (q^0, \val{0}{})$ and $\goesto{}$ is the minimal transition
  relation satisfying the rule
  %
  \begin{equation}
    \label{eq:comp:semantics}
    \infer{
      q \goesto{a, g, u} q'
      \and
      \val{}{} \models g
      \and
      \val[\primeit]{}{} = \tilde{\val{}{}}[u]
      \and
      \valdiff{\val{}{}}{\tilde{\val{}{}}} \subseteq \export(a)
    }{
      (q, \val{}{}) \goesto{a,\tilde{\val{}{}}} (q', \val[\primeit]{}{})
    }
    \,.
  \end{equation}
  %
  \todoSB{Do we actually need this?}\todoLH{I agree: I do not think we need this}
  \hl{The \emph{closed semantics} of $B$ is given by the LTS denoted
  $\semclosed{B} \bydef{=}\reachable{\semopen{B}}$, comprising
  only the reachable states of $\semopen{B}$.}
\end{definition}

@@

  An \emph{interaction model over $(V, P, \export)$} is a set $\Gamma
  \subseteq 2^P \times \guards{V} \times \updates{V}$, such that,
  for any interaction $(a, g, u) \in \Gamma$, we have
  $g \in \guards{\export(a)}$ and $e \in \updates{\export(a)}$.\footnote{%
%
    Notice that this definition allows $(\emptyset, \true,
    \noop)$ and $(\emptyset, \false, \noop)$ to be included in
    $\Gamma$.
%
  }
  %% We call the set of ports $P$ the \emph{domain} of the
  %% interaction model.
\end{definition}

%% \noteLH{supp should go outside the definition}
When speaking of a set of components or interfaces, we will always
assume that it satisfies the disjointness assumption of \defn{im}.
%
We call the \emph{support} of a set of ports $a \subseteq P$, denoted
$\supp{a}$, either the set $\setdef{i \in I}{a \cap P_i \neq
  \emptyset}$ (for $P = \bigcup_{i=1}^n P_i$) or the set $\setdef{B
  \in \cB}{a \cap P_B \neq \emptyset}$ (for $P = \bigcup_{B \in \cB}
P_B$).  The precise meaning of this notation will always be clear from
the context.  Intuitively, for an interaction $(a, g, u)$, $\supp{a}$
denotes the set of the participating components.

\begin{definition}[Composition]
  \label{defn:im:sem}
  The \emph{composition} of a finite set of components $\cB = (Q_i,
  q^0_i, V_i, \val{0}{i}, P_i, \export[i], \goesto{})^{i \in I}$ with
  the interaction model $\Gamma$ over $(V, P, \export)$ (as in \defn{im})
  is the component $\Gamma(\cB) = (Q, q^0, V, \val{0}{}, P,
    \export, \goesto{})$, where
%
  $Q = \prod_{i \in I} Q_i$;
%
  $q^0 = (q_i^0)^{i \in I}$;
%
  $\val{0}{}: V \rightarrow \data$ is such that, for any $v \in
  V_i$, $\val{0}{}(v) = \val{0}{i}(v)$;
%
  and $\goesto{}$ is the minimal transition relation satisfying
  the rule
%
  \begin{gather}
%%     \label{eq:im:empty}
%%     \infer{
%%       q_j \goesto{\emptyset, g, e} q_j'
%%       \and
%%       \forall i \neq j, q_i = q_i'
%%     }{
%%       (q_i)^{i \in I} \goesto{\emptyset, g, e} (q_i')^{i \in I}
%%     }\,,
%% %
%%     \\[0.5\baselineskip]
    \label{eq:im:sem}
%
    \infer{    
      (a, g, u) \in \Gamma
      \and
      a \neq \emptyset
      \and
      \forall i \in \supp{a}, q_i \goesto{a \cap P_i, g_i, u_i} q_i'
      \and
      \forall i \not\in \supp{a}, q_i = q_i'
%
      \\\\
      \textstyle
%
      g' = g \land \bigwedge_{i \in \supp{a}} g_i
      \and
      u' = u; u_i^{i \in \supp{a}}
    }{
      (q_i)^{i \in I} \goesto{a, g', u'} (q_i')^{i \in I}
    }\,.
  \end{gather}
  \todoSBin{Add the intuitive explanation}
\end{definition}

@@

For instance, in \fig{schema:ArchFailure:BIP}, the two ports
$\NameTimer.\PortStart$ and $\NameCtrl.\PortFail$ are always
synchronised, since they belong to exactly one binary sub-connector,
where they are both synchrons.  In particular, this means that
whenever the transition $\cstate{0} \goesto{\PortFail} \cstate{1}$ is
fired, so is the transition $\tstate{0} \goesto{\PortStart,
  \TimerInit[]} \tstate{1}$, initialising the timer.  The binary
connector $\NameTimer.\PortStart \bullet\!\!\!-\!\!\!-\!\!\!\bullet
\NameCtrl.\PortFail$ is a sub-connector of a hierarchical connector,
where the port $\NameOprnd.\PortFail$ is a trigger.  Thus, the above
interaction can only happen together with $\NameOprnd.\PortFail$,
forming a ternary interaction.  On the contrary, being a trigger, the
port $\NameOprnd.\PortFail$ can fire alone, forming a singleton
interaction.
